---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- run rmarkdown::render('README.Rmd', output_file = 'README.md', encoding = 'utf8') -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# kaefa

The goal of kaefa is to improving research capability to identify unexplained factor structure with complexly cross-classified multilevel structured data in R environment with automated exploratory factor analysis (aefa) framework

## Algorithm

The automated exploratory factor analysis (aefa) framework implements a **greedy search algorithm** to find the global-optimal model configuration. The algorithm iteratively:

1. Evaluates multiple model candidates with different factor structures and item response models
2. Selects the best model based on information criteria (DIC, AIC, BIC, etc.)
3. Assesses item fit and removes poorly fitting items one at a time
4. Re-estimates the model until convergence to an optimal solution

This greedy approach enables efficient exploration of the model space while seeking globally-optimal solutions. The method is inspired by research on globally-optimal greedy algorithms in tracking and optimization problems (Pirsiavash, Ramanan, & Fowlkes, 2011).

**References:**

- Pirsiavash, H., Ramanan, D., & Fowlkes, C. C. (2011, June). Globally-optimal greedy algorithms for tracking a variable number of objects. In Computer Vision and Pattern Recognition (CVPR), 2011 IEEE Conference on (pp. 1201-1208). IEEE.
- Wikipedia: [Greedy Algorithm (탐욕 알고리즘)](https://ko.m.wikipedia.org/wiki/탐욕_알고리즘)

## Installation

You can install kaefa from github with:

```{r gh-installation, eval = TRUE}
# install.packages("devtools")
devtools::install_github("seonghobae/kaefa")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
library('kaefa')
mod1 <- kaefa::aefa(mirt::Science)
mod1
```

## software quality information

### ubuntu and mac environment
[![Travis-CI Build Status](https://travis-ci.org/seonghobae/kaefa.svg?branch=master)](https://travis-ci.org/seonghobae/kaefa)

### windows environment
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/seonghobae/kaefa?branch=master&svg=true)](https://ci.appveyor.com/project/seonghobae/kaefa)

<!-- ### code quality -->
<!-- [![Coverage Status](https://img.shields.io/codecov/c/github/seonghobae/kaefa/master.svg?maxAge=3600)](https://codecov.io/github/seonghobae/kaefa?branch=master) -->

[Contributor Code of Conduct](CONDUCT.md)
