---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- run rmarkdown::render('README.Rmd', output_file = 'README.md', encoding = 'utf8') -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# kaefa

The goal of kaefa is to improving research capability to identify unexplained factor structure with complexly cross-classified multilevel structured data in R environment with automated exploratory factor analysis (aefa) framework

## Algorithm

The automated exploratory factor analysis (aefa) framework implements a **greedy search algorithm** to efficiently explore the model space and find improved model configurations. The algorithm iteratively:

1. Evaluates multiple model candidates with different factor structures and item response models
2. Selects the best model based on information criteria (DIC, AIC, BIC, etc.)
3. Assesses item fit and removes poorly fitting items one at a time
4. Re-estimates the model until convergence to a locally optimal solution

This greedy approach enables efficient exploration of the model space while seeking improved solutions through iterative refinement. The method aligns with model selection and exploratory factor analysis research (Preacher, Zhang, Kim, & Mels, 2013; Jennrich & Bentler, 2011).

**References:**

- Preacher, K. J., Zhang, G., Kim, C., & Mels, G. (2013). Choosing the optimal number of factors in exploratory factor analysis: A model selection perspective. Multivariate Behavioral Research, 48(1), 28-56. https://doi.org/10.1080/00273171.2012.710386
- Jennrich, R. I., & Bentler, P. M. (2011). Exploratory bi-factor analysis. Psychometrika, 76(4), 537-549. https://doi.org/10.1007/s11336-011-9218-4

## Installation

You can install kaefa from github with:

```{r gh-installation, eval = TRUE}
# install.packages("devtools")
devtools::install_github("seonghobae/kaefa")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
library('kaefa')
mod1 <- kaefa::aefa(mirt::Science)
mod1
```

## software quality information

### ubuntu and mac environment
[![Travis-CI Build Status](https://travis-ci.org/seonghobae/kaefa.svg?branch=master)](https://travis-ci.org/seonghobae/kaefa)

### windows environment
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/seonghobae/kaefa?branch=master&svg=true)](https://ci.appveyor.com/project/seonghobae/kaefa)

<!-- ### code quality -->
<!-- [![Coverage Status](https://img.shields.io/codecov/c/github/seonghobae/kaefa/master.svg?maxAge=3600)](https://codecov.io/github/seonghobae/kaefa?branch=master) -->

[Contributor Code of Conduct](CONDUCT.md)
