---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- run rmarkdown::render('README.Rmd', output_file = 'README.md', encoding = 'utf8') -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# kaefa

[![Ask DeepWiki](https://deepwiki.com/badge.svg)](https://deepwiki.com/ContextualWisdomLab/kaefa)

**Automated exploratory IRT/factor-model search for complex assessment data in
R.**

kaefa helps researchers explore uncertain factor structures without manually
rebuilding every candidate model. Its AEFA workflow coordinates candidate
estimation, information-criterion comparison, item-fit review, and iterative
refinement while keeping the fitted `mirt` model artifacts available for
scientific inspection.

## Where kaefa fits

Use kaefa when the factor structure is not fully known in advance and you need a
repeatable exploratory search over supported IRT/factor-model candidates. The
package is aimed at psychometric and applied measurement work, including data
with multilevel or cross-classified structure where supported by the underlying
model path.

kaefa does not turn an exploratory search into confirmatory evidence. Factor
interpretation, model adequacy, recovery, sampling assumptions, and downstream
score-use decisions still require domain review and appropriate validation.

## Core workflow

1. Provide an item-response data set or supported model inputs.
2. Run `aefa()` with the candidate/search settings appropriate to the study.
3. kaefa evaluates candidate models through `engineAEFA()` and the configured
   fit/search policy.
4. Review the selected model, item-fit evidence, convergence diagnostics, and
   retained search history before interpreting the result.

A minimal scientific API example is:

```{r example, eval = FALSE}
library(kaefa)

result <- kaefa::aefa(mirt::Science)
result
```

The current numerical estimation boundary is `mirt`; kaefa orchestrates model
search and supporting diagnostics rather than owning an independent IRT
likelihood engine.

## Model-search behavior

The AEFA loop uses a greedy candidate-search strategy. AIC is the default
selection criterion, with AICc, BIC, and sample-size-adjusted BIC available where
the fitted model supplies the required quantities. DIC is used only when a
fitted model actually supplies posterior DIC; it is not approximated with AIC.

Item-fit/search behavior and accepted `mirt` fit-search decisions are documented
in [`docs/adr/`](docs/adr/). The research/provenance index is maintained in
[`docs/papers/README.md`](docs/papers/README.md).

## Product surfaces

kaefa currently remains one R package with three explicit product boundaries:

- **kaefa-core** — `aefa()`, `engineAEFA()`, model selection, item-fit evaluation,
  theta-prior utilities, and scientific/recovery evidence.
- **kaefa-studio** — the optional bundled Shiny interface launched with
  `launchAEFA()`.
- **kaefa-runner** — the future hosted/container/remote execution boundary; it
  is not a separately released product today.

Optional remote execution is configured through `aefaInit()`. Local execution
remains the default. SSH keys and other credentials must stay outside the
repository and be supplied through appropriately protected local/hosted
configuration.

## Evaluate the source

The repository does not currently publish a GitHub Release. Source metadata is
`0.1.428`; that value and a passing development check are not immutable release
or support evidence.

For maintainers working in an isolated dependency/license-review environment,
run the repository package checks against the checkout rather than relying on
historical `seonghobae/kaefa`, Travis CI, or AppVeyor instructions:

```bash
R CMD build .
R CMD check --no-manual kaefa_*.tar.gz
```

The exact supported R/dependency environment and hosted evidence are defined by
current repository workflows and package metadata, not by historical badges in
this README.

## Interactive studio

The Shiny surface exposes the same underlying package responsibility for users
who prefer an interactive workflow:

```{r shiny-launch, eval = FALSE}
library(kaefa)
launchAEFA()
```

Treat UI output as analysis evidence to review, not as an automatic scientific
or operational decision.

## Architecture and verification

The current runtime flow is:

`aefa()` -> candidate search -> `engineAEFA()` / `mirt` estimation -> model
comparison -> item-fit/refinement -> selected model + diagnostics.

See [ARCHITECTURE.md](ARCHITECTURE.md) for repository layout, runtime flow,
remote-execution boundaries, and quality gates. Scientific changes should carry
focused regression/recovery evidence; documentation-only checks do not prove
parameter recovery or model validity.

## Commercial licensing status

**Not currently cleared for ContextualWisdomLab commercial
intake/distribution.**

`DESCRIPTION` declares kaefa itself as `GPL-3` and directly imports `mirt`, a
GPL-family runtime dependency, plus additional R dependencies that require a
complete current license inventory. There is no root Apache-2.0/MIT `LICENSE`
file that overrides those terms.

Issue #83 owns the first-party provenance/relicensing review and replacement of
the GPL-family runtime/build dependency boundary while preserving intentionally
supported scientific behavior. Until that work is integrated and verified, do
not present kaefa as Apache-2.0/MIT-cleared or commercially policy-compliant.

## Documentation and contribution

- [Architecture](ARCHITECTURE.md) — product/runtime ownership boundaries.
- [Architecture decisions](docs/adr/README.md) — accepted fit/search decisions.
- [Research sources](docs/papers/README.md) — formula and literature provenance.
- [Contributor Code of Conduct](CONDUCT.md) — collaboration expectations.
- [Ask DeepWiki](https://deepwiki.com/ContextualWisdomLab/kaefa) —
  repository-aware navigation.

For scientific behavior changes, include the estimand/model assumption being
changed and reproducible recovery/regression evidence. For documentation or
operability changes, keep user-facing claims tied to current protected source
and exact-head checks.
