---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- run rmarkdown::render('README.Rmd', output_file = 'README.md', encoding = 'utf8') -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# kaefa

The goal of kaefa is to improving research capability to identify unexplained factor structure with complexly cross-classified multilevel structured data in R environment with automated exploratory factor analysis (aefa) framework

## Installation

You can install kaefa from github with:

```{r gh-installation, eval = TRUE}
# install.packages("devtools")
devtools::install_github("seonghobae/kaefa")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
library('kaefa')
mod1 <- kaefa::aefa(mirt::Science)
mod1
```

## New Feature: fitdistrplus Integration for Theta Priors

kaefa now supports setting theta priors based on empirical raw score distributions using the `fitdistrplus` package. This feature allows you to:

1. Fit distributions to raw scores to inform theta priors
2. Test if calibration works for non-nominal models
3. Validate model calibration against empirical distributions

Example usage:

```{r fitdistrplus-example, eval = FALSE}
# Fit a distribution to raw scores
fit <- fitThetaPrior(mirt::Science, dist = "norm")

# Test calibration with distribution fit
testResult <- testThetaPriorCalibration(mirt::Science, dist = "norm")

# Apply theta prior during calibration
model <- applyThetaPrior(mirt::Science, fit, minExtraction = 1, maxExtraction = 1)
```

For more examples and detailed documentation, see the `examples/` directory.

## software quality information

### ubuntu and mac environment
[![Travis-CI Build Status](https://travis-ci.org/seonghobae/kaefa.svg?branch=master)](https://travis-ci.org/seonghobae/kaefa)

### windows environment
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/seonghobae/kaefa?branch=master&svg=true)](https://ci.appveyor.com/project/seonghobae/kaefa)

<!-- ### code quality -->
<!-- [![Coverage Status](https://img.shields.io/codecov/c/github/seonghobae/kaefa/master.svg?maxAge=3600)](https://codecov.io/github/seonghobae/kaefa?branch=master) -->

[Contributor Code of Conduct](CONDUCT.md)
